---
title: "Network_analysis"
author: "Abdelrahman"
date: "2023-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task detailes

• The assignment is based on this paper: https://www.pnas.org/doi/10.1073/pnas.0701361104
  (skimming over the paper is sufficient; the specific research topic is not of any particular     interest to us)

• Load the Human Diseasome (Diseasome_preprocessed.Rdata) into the workspace. Familiarize
  yourself with the three data frames. The data frame diseasome is the Human Diseasome (paper    Fig.1, middle) in edge list format. The first two columns are indices into the disease and     gene list,respectively. How many diseases and genes are in the Diseasome? How many edges do    you see?

• Transform the loaded edge list into a matrix, where rows correspond to diseases, columns to    genes,and an entry to whether an edge exists between the respective disease and gene. What     are the dimensions of the matrix?

• Generate the Human Disease Network (paper Fig. 1, left). That is, create a symmetric           disease-bydisease adjacency matrix where each cell contains the number of genes those two      diseases share. You do not need to generate a network figure; the matrix is enough.

• How many total (unweighted) edges does the network have?


### How many diseases and genes are in the Diseasome? How many edges do you see?

```{r}
load("Diseasome_preprocessed.RData")

#explore the diseasome data frame
str(diseasome)

# number of unique diseases 
num_unique_diseases <- length(unique(diseasome$Disease_ID))
num_unique_diseases
# number of unique genes 
num_unique_genes <- length(unique(diseasome$Gene_ID))
num_unique_genes
#number of edges 
length(diseasome$Edge)
```

### What are the dimensions of the matrix?

```{r}
#matrix of the diseasome 
diseasome_matrix <- matrix(nrow = length(diseasome$Disease_ID) , ncol=length(diseasome$Gene_ID) )

dim(diseasome_matrix)

```

## Note: because of the limited computational resources I just picked the first 100 entries in the Diseasome data set BUT this code applicable for the all datasets in case of availability of this resource 
```{r}

diseasome_matrix <- matrix(0,nrow = 100 , ncol =100)

#fill the matrix by the edges 
for(i in 1:nrow(diseasome[c(1:100),])){
  diseasome_matrix[diseasome[i,1],diseasome[i,2]] <- 1
}


#dimensions of the matrix 
dim(diseasome_matrix)

```

### create a symmetric disease-by-disease adjacency matrix where each cell contains the number of genes those two diseases share

```{r}

# Transpose the diseasome_matrix
gene_disease_matrix <- t(diseasome_matrix)
gene_disease_matrix
# Multiply the transposed matrix by the original matrix
gene_gene_matrix <- gene_disease_matrix %*% diseasome_matrix
gene_gene_matrix
# Convert the gene-by-gene matrix to a disease-by-disease matrix
disease_disease_matrix <- t(gene_gene_matrix) %*% gene_gene_matrix

#give the rows and columns names 
rownames(disease_disease_matrix) <- colnames(disease_disease_matrix) <- paste0("D", 1:nrow(disease_disease_matrix))

disease_disease_matrix


```

### How many total (unweighted) edges does the network have?

```{r}
# Count the number of edges THEN 
#subtract the number of diagonal values which represented by the number of rows 
num_edges <- sum(disease_disease_matrix > 0) - nrow(disease_disease_matrix)

# Multiply by 2 to account for symmetry
total_edges <- 2 * num_edges

total_edges
```



